Hydrodynamics and rheology of a vesicle doublet suspension

### 2019-06-17 vesicle_code/docs/adhesion/adhesion.tex
working on revisions

\usepackage{amsmath,amssymb}
\usepackage[titletoc]{appendix}
\usepackage{todonotes}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{stmaryrd}
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

@@

\begin{abstract}
The dynamics of an adhesive vesicle doublet under various flow conditions is
investigated numerically using a high-order, adaptive-in-time boundary
integral method. In a quiescent flow, two nearby vesicles
move slowly towards each other under the adhesive potential, pushing out
fluid between them to form a vesicle doublet at equilibrium. A lubrication
analysis on such draining of a thin film gives the dependencies of
draining time on adhesion strength and separation distance that are in
good agreement with numerical results.
%
In a microfluid trap where the stagnation of an extensional flow is dynamically
placed in the middle of a vesicle doublet through an active control
loop, novel dynamics of a vesicle doublet is observed. Numerical
simulations show that there exists a critical extensional flow rate
above which adhesive interaction is overcome by the 
diverging stream, thus providing a simple method to measure the adhesion
strength between two vesicle membranes. 
%
In a planar shear flow, numerical simulations reveal that a vesicle
doublet may form provided that the adhesion
strength is sufficiently large at a given vesicle reduced area. Once a
doublet is formed, the effective shear viscosity of a dilute suspension
of vesicle doublets is found to be a function of  reduced area.  Results from these numerical studies and
analysis shed light on the hydrodynamic and rheological consequences of adhesive interactions
between vesicles in a viscous fluid.
%\begin{description}
%\item[Usage]
%Secondary publications and information retrieval purposes.
%\item[PACS numbers]
%May be entered using the \verb+\pacs{#1}+ command.
%\item[Structure]
%You may use the \texttt{description} environment to structure your abstract;
%use the optional argument of the \verb+\item+ command to give the category of each item. 
%\end{description}
\end{abstract}

@@

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by the reduced volume (reduced area in two dimensions),
viscosity contrast between interior and exterior fluids, and shear rate
of the imposed far-field fluid flow. In addition, a vesicle with a rigid
particle inside is also investigated as a biological mimic of a
eukaryotic cell with a nucleus that occupies nearly half of the
intracellular volume~\cite{Veerapaneni2011_PRL}. Small-deformation
analysis shows that a vesicle tank-treads in a planar shear flow for low
viscosity contrast and shear rate. At high viscosity contrast the
tank-treading dynamics transitions to tumbling
dynamics~\cite{Misbah2006_PRL, Vlahovska2007_PRE}, and this leads to a
transition in the effective shear viscosity of the vesicle
suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also validated
by direct numerical simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{DeschampsKantsler2009_PNAS,
KantslerSegreSteinberg2008_EPL, ZabuskySegreDeschamps2011_PoF}.  Between
tank-treading and tumbling vesicle dynamics, a breathing (tremble) mode
is also observed~\cite{Misbah2006_PRL, KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2011_JFM, SpannZhaoShaqfeh2014_PoF} to alter the effective
shear viscosity.  In an extensional flow (planar or uniaxial), vesicle
shape dynamics depends sensitively on the vesicle reduced volume and the
elastic capillary number~\cite{KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2013_JFM, Narsimhan2014_JFM,
DahlNarsimhanGouveia2016_SoftMatt}: Asymmetric shape and oscillatory
undulation of the vesicle membrane are two examples of the complex
vesicle hydrodynamics in an extensional flow.

@@

We characterize the effect of the Hamaker constant and shear rate by
determining a critical Hamaker constant, $\mathcal{H}_c$, which
determines if the vesicles form a doublet or separate.  We fix the
separation distance to $\delta = 0.4$ and the vesicle reduced area to
$\Delta A = 0.9$.  In the left plot of
Figure~\ref{fig:sflow_phase_diagram}, we plot the minimum distance
between the vesicles as a function of time for various Hamaker
constants.  The red curves correspond to Hamaker constants that are too
small to bind the vesicles into a doublet, and the blue curves
correspond to Hamaker constants that are sufficiently large to form a
doublet.

\begin{figure}
  \includegraphics[height=0.27\textwidth]{figs/shear_adR4em1Chi5em1_ra090.pdf}
  \includegraphics[height=0.27\textwidth]{figs/shear_adR4em1_ra090_phaseDiagram.pdf}
  \includegraphics[height=0.27\textwidth]{figs/shear_adR1em1_ra090_phaseDiagram.pdf}
  \caption{\label{fig:sflow_phase_diagram} {\em Left}: Distance between
  a pair of vesicles in a planar shear flow with shear rate $\chi=0.5$,
  and separation distance $\delta = 0.4$.  The different lines
  correspond to a linear spacing of Hamaker constants ranging from
  $\mathcal{H}=0.1$ and $\mathcal{H}=1.0$.  {\em Center and Right}:
  Phase diagram of vesicle hydrodynamics in a shear flow with $\delta =
  0.4$ (center) and $\delta = 0.1$ (right) with reduced area  $\Delta A
  = 0.9$. The critical Hamaker constant $\mathcal{H}$ for binding of two
  vesicles in a shear flow depends on the shear rate $\chi$.  For a
  doublet to form, lower separation distances require a stronger Hamaker
  constants at a given shear rate.}
\end{figure}

@@

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.45\textwidth]{figs/shear2Ves_adR4em1adS7em1Chi5em1.pdf}
  \includegraphics[width=0.45\textwidth]{figs/doublet_decomp.pdf}
  \caption{\label{fig:shearIntrinsicViscosity} {\em Left}: The intrinsic
  viscosity of a tank-treading vesicle (blue) and a doublet (red).  The
  shear rate is $\chi = 0.5$, the Hamaker constant is $\mathcal{H} =
  0.7$, and the separation distance is $\delta = 0.4$.  The black marks
  denote intrinsic viscosity values computed by Ghigliotti {\em et
  al.}~\cite{GhigliottiBibenMisbah2010_JFM} (cf.~Figure 5).  {\em
  Right}: The decomposition of the intrinsic viscosity into the
  contributions for the bending and tension (blue) of the vesicles, and
  the contribution from the vesicle adhesion (red).  Also included is
  the intrinsic viscosity of a single tank-treading vesicle (dashed
  black line).}
\end{figure}

>>>>

\usepackage{amsmath,amssymb}
\usepackage[titletoc]{appendix}
\usepackage{todonotes}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{stmaryrd}
%\usepackage{subfig}
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

\newcommand{\subfigimg}[3][,]{%
  \setbox1=\hbox{\includegraphics[#1]{#3}}% Store image in box
  \leavevmode\rlap{\usebox1}% Print image
  \rlap{\hspace*{0pt}\raisebox{\dimexpr\ht1-0\baselineskip}{\bf #2}}% Print label
  \phantom{\usebox1}% Insert appropriate spcing
}

@@

\begin{abstract}
The dynamics of an adhesive two-dimensional vesicle doublet under
various flow conditions is investigated numerically using a high-order,
adaptive-in-time boundary integral method. In a quiescent flow, two
nearby vesicles move slowly towards each other under the adhesive
potential, pushing out fluid between them to form a vesicle doublet at
equilibrium. A lubrication analysis on such draining of a thin film
gives the dependencies of draining time on adhesion strength and
separation distance that are in good agreement with numerical results.
%
In a microfluid trap where the stagnation of an extensional flow is
dynamically placed in the middle of a vesicle doublet through an active
control loop, novel dynamics of a vesicle doublet is observed. Numerical
simulations show that there exists a critical extensional flow rate
above which adhesive interaction is overcome by the diverging stream,
thus providing a simple method to measure the adhesion strength between
two vesicle membranes. 
%
In a planar shear flow, numerical simulations reveal that a vesicle
doublet may form provided that the adhesion strength is sufficiently
large at a given vesicle reduced area. Once a doublet is formed, the
effective shear viscosity of a dilute suspension of vesicle doublets is
found to be a function of  reduced area.  Results from these numerical
studies and analysis shed light on the hydrodynamic and rheological
consequences of adhesive interactions between vesicles in a viscous
fluid.
%\begin{description}
%\item[Usage]
%Secondary publications and information retrieval purposes.
%\item[PACS numbers]
%May be entered using the \verb+\pacs{#1}+ command.
%\item[Structure]
%You may use the \texttt{description} environment to structure your abstract;
%use the optional argument of the \verb+\item+ command to give the category of each item. 
%\end{description}
\end{abstract}

@@

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by the reduced volume (reduced area in two dimensions),
viscosity contrast between interior and exterior fluids, and shear rate
of the imposed far-field fluid flow. In addition, a vesicle with a rigid
particle inside is also investigated as a biological mimic of a
eukaryotic cell with a nucleus that occupies nearly half of the
intracellular volume~\cite{Veerapaneni2011_PRL}. Small-deformation
analysis shows that a vesicle tank-treads in a planar shear flow for low
viscosity contrast and shear rate. At high viscosity contrast the
tank-treading dynamics transitions to tumbling
dynamics~\cite{Misbah2006_PRL, Vlahovska2007_PRE}, and this leads to a
transition in the effective shear viscosity of the vesicle
suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also validated
by direct numerical simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{Vitkova2008_BJ,DeschampsKantsler2009_PNAS,
KantslerSegreSteinberg2008_EPL, ZabuskySegreDeschamps2011_PoF}.  Between
tank-treading and tumbling vesicle dynamics, a breathing (tremble) mode
is also observed~\cite{Misbah2006_PRL, KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2011_JFM, SpannZhaoShaqfeh2014_PoF} to alter the effective
shear viscosity.  In an extensional flow (planar or uniaxial), vesicle
shape dynamics depends sensitively on the vesicle reduced volume and the
elastic capillary number~\cite{KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2013_JFM, Narsimhan2014_JFM,
DahlNarsimhanGouveia2016_SoftMatt}: Asymmetric shape and oscillatory
undulation of the vesicle membrane are two examples of the complex
vesicle hydrodynamics in an extensional flow.

@@

We characterize the effect of the Hamaker constant and shear rate by
determining a critical Hamaker constant, $\mathcal{H}_c$, which
determines if the vesicles form a doublet or separate.  We fix the
separation distance to $\delta = 0.4$ and the vesicle reduced area to
$\Delta A = 0.9$.  In the left plot of
Figure~\ref{fig:sflow_distance}, we plot the minimum distance
between the vesicles as a function of time for various Hamaker
constants.  The red curves correspond to Hamaker constants that are too
small to bind the vesicles into a doublet, and the blue curves
correspond to Hamaker constants that are sufficiently large to form a
doublet.  Figure~\ref{fig:sflow_phase_diagram} indicates whether a
doublet with reduced area $\Delta A = 0.9$ is formed (blue) or not (red)
for a variety of shear rates, Hamaker constants, and at two different
separation distances.  The plot includes a line of best fit between the
stable and unstable regions.  Since the transition is linear, this
indicates that a dimensionless number involving the ratio of the Hamaker
constant could be used to determine if a doublet forms or not.

\begin{figure}
  \includegraphics[height=0.27\textwidth]{figs/shear_adR4em1Chi5em1_ra090.pdf}
  \caption{\label{fig:sflow_distance} {\em Left}: The distance between a
  pair of vesicles in a planar shear flow with shear rate $\chi=0.5$,
  and separation distance $\delta = 0.4$.  The different lines
  correspond to a linear spacing of Hamaker constants ranging from
  $\mathcal{H}=0.1$ and $\mathcal{H}=1.0$.  {\em Right}:
  \todo[inline]{frequency of the oscillation}}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[height=0.37\textwidth]{figs/shear_adR1em1_ra090_phaseDiagram.pdf}
  \includegraphics[height=0.37\textwidth]{figs/shear_adR4em1_ra090_phaseDiagram.pdf}
  \caption{\label{fig:sflow_phase_diagram} Phase diagram of vesicle
  hydrodynamics in a shear flow with $\delta = 0.1$ (left) and $\delta =
  0.4$ (right) with reduced area $\Delta A = 0.9$. The critical Hamaker
  constant $\mathcal{H}$ for binding of two vesicles in a shear flow
  depends on the shear rate $\chi$.  For a doublet to form, lower
  separation distances require a stronger Hamaker constants at a given
  shear rate.  The stability limit appears to be linear, indicating that
  there is potentially a stability criteria that depends on a
  dimensionless number involving the ratio of the Hamaker constant and
  shear rate.}
\end{figure}

@@

\begin{figure}[]
  \begin{tabular}{@{}p{0.45\linewidth}@{\quad}p{0.45\linewidth}@{}}
  \subfigimg[width=\linewidth]{(a)}{figs/doublet_decomp.pdf} &
  \subfigimg[width=\linewidth]{(b)}{figs/doublet_decomp.pdf}
  \end{tabular}
%  \centering
%  \includegraphics[width=0.45\textwidth]{figs/shear2Ves_adR4em1adS7em1Chi5em1.pdf}
%  \includegraphics[width=0.45\textwidth]{figs/doublet_decomp.pdf}
  \caption{\label{fig:shearIntrinsicViscosity} {\em Left}: The intrinsic
  viscosity of a tank-treading vesicle (blue) and a doublet (red).  The
  shear rate is $\chi = 0.5$, the Hamaker constant is $\mathcal{H} =
  0.7$, and the separation distance is $\delta = 0.4$.  The black marks
  denote intrinsic viscosity values computed by Ghigliotti {\em et
  al.}~\cite{GhigliottiBibenMisbah2010_JFM} (cf.~Figure 5).  {\em
  Right}: The decomposition of the intrinsic viscosity into the
  contributions for the bending and tension (blue) of the vesicles, and
  the contribution from the vesicle adhesion (red).  Also included is
  the intrinsic viscosity of a single tank-treading vesicle (dashed
  black line).}
\end{figure}

### 2019-06-17 vesicle_code/docs/adhesion/revisions/reviewer2.tex
starting revisions on adhesion paper
working on revisions

\usepackage{fullpage}
\usepackage{todonotes}
\usepackage{amsmath,amsfonts,amssymb,stmaryrd}
\usepackage{color}
\newcommand{\comment}[1]{{\color{blue} #1}}

@@

\noindent
\comment{Abstract: it should be mentioned in the abstract that this is 2D
modelling. 2D simulations of vesicles are often perfectly fine to
investigate the physics of phenomena but in some cases 3D effects may
have their importance, at least for quantitative comparisons with
experiments. The reader should identify either from the title or
abstract that this study is 2D.}
\begin{itemize}
  \item rebuttal
\end{itemize}

\noindent
\comment{Introduction, 2nd paragraph: the transition in the effective
viscosity of vesicle or RBC suspensions was also validated
experimentally in ref [18] (Vitkova et al.), not only in [20-22].}
\begin{itemize}
  \item rebuttal
\end{itemize}

@@

\noindent
\comment{Fig 9(right): I regret that the authors did not extend the study
to lower values of the reduced area, to reach values comparable to red
blood cells which are more deflated that $\Delta A=0.7$. Also, the range
of shear rates could have been slightly increased to reach the
stability limit of the most deflated vesicles. Here, we do not know
what happens below reduced areas of 0.8.}
\begin{itemize}
  \item rebuttal
\end{itemize}

@@

\noindent
\comment{Fig 13 and comments about rheology: I am not sure to understand
the meaning of the left figure where it is shown that the intrinsic
viscosity of the doublet in suspension is about twice the viscosity of a
single vesicle. By definition, the intrinsic viscosity does not depend
on volume fraction, so I do not expect such a big difference between
bound vesicles and isolated vesicles. With the values chosen for the
Hamaker constant and the shear rate, it seems (from fig 12-left) that
the interaction is quite strong and that the distance between the two
vesicles in the doublet is always very short (close to the equilibrium
distance $\delta$). So the doublet must be roughly tumbling as a single
object and unless a co-rotating tank-treading motion of the two
membranes produces a strong dissipation by shear flow in the gap between
the two vesicles, the difference should not be so large. It is also
quite curious that the intrinsic viscosity of the doublet is so close to
exactly twice the viscosity of a single vesicle. Could the authors check
and elaborate on this ? Was the volume fraction properly scaled out
here? Also in the right panel, the bending/tension part is compared to
twice the intrinsic viscosity of a suspension of isolated vesicles. Why
twice ? Finally, it is not clear how technically the bending/tension and
adhesion parts are separated.  Overall I think the study on rheology
needs much more development than what is presented here with only one
set of parameter values and very little quantitative comments.}
\begin{itemize}
  \item rebuttal
\end{itemize}

>>>>

\usepackage{fullpage}
\usepackage{todonotes}
\usepackage{amsmath,amsfonts,amssymb,stmaryrd}
\usepackage{color}
\newcommand{\comment}[1]{{\color{blue} #1}}
\newcommand{\xxi}{\boldsymbol{\xi}}

@@

\noindent
\comment{Abstract: it should be mentioned in the abstract that this is
2D modelling. 2D simulations of vesicles are often perfectly fine to
investigate the physics of phenomena but in some cases 3D effects may
have their importance, at least for quantitative comparisons with
experiments. The reader should identify either from the title or
abstract that this study is 2D.}
\begin{itemize}
  \item We now mention in the abstract that we consider the dynamics of
    and adhesive two-dimensional vesicle doublet.
\end{itemize}

\noindent
\comment{Introduction, 2nd paragraph: the transition in the effective
viscosity of vesicle or RBC suspensions was also validated
experimentally in ref [18] (Vitkova et al.), not only in [20-22].}
\begin{itemize}
  \item We have included reference [18] in the list of experimental
    works that validated the transition in the effective viscosity
\end{itemize}

@@

\noindent
\comment{Fig 9(right): I regret that the authors did not extend the study
to lower values of the reduced area, to reach values comparable to red
blood cells which are more deflated that $\Delta A=0.7$. Also, the range
of shear rates could have been slightly increased to reach the
stability limit of the most deflated vesicles. Here, we do not know
what happens below reduced areas of 0.8.}
\begin{itemize}
  \item We have extended the results to include vesicles of reduced area
    0.65, a typical value for two-dimensional red blood cells.
  \item We have also reported results for larger extensional rates so
    that the transition from a fluid trap to two separating vesicles can
    be observed at all the reduced areas.
\end{itemize}

@@

\noindent
\comment{Fig 13 and comments about rheology: I am not sure to understand
the meaning of the left figure where it is shown that the intrinsic
viscosity of the doublet in suspension is about twice the viscosity of a
single vesicle. By definition, the intrinsic viscosity does not depend
on volume fraction, so I do not expect such a big difference between
bound vesicles and isolated vesicles. With the values chosen for the
Hamaker constant and the shear rate, it seems (from fig 12-left) that
the interaction is quite strong and that the distance between the two
vesicles in the doublet is always very short (close to the equilibrium
distance $\delta$). So the doublet must be roughly tumbling as a single
object and unless a co-rotating tank-treading motion of the two
membranes produces a strong dissipation by shear flow in the gap between
the two vesicles, the difference should not be so large. It is also
quite curious that the intrinsic viscosity of the doublet is so close to
exactly twice the viscosity of a single vesicle. Could the authors check
and elaborate on this ? Was the volume fraction properly scaled out
here? Also in the right panel, the bending/tension part is compared to
twice the intrinsic viscosity of a suspension of isolated vesicles. Why
twice ? Finally, it is not clear how technically the bending/tension and
adhesion parts are separated.  Overall I think the study on rheology
needs much more development than what is presented here with only one
set of parameter values and very little quantitative comments.}
\begin{itemize}
  \item The reviewer is correct that we miscalculated the intrinsic
    viscosity.  We had initially divided by only the area inside a
    single vesicle rather than both vesicles in the doublet.  This has
    been corrected, and now there is only a small increase in the
    intrinsic viscosity when compared to a single tank treading vesicle.

  \item With the numbers corrected, the intrinsic viscosity is clearly
    not equal to the intrinsic viscosity of a tank treading vesicle.

  \item We have double-checked our calculations and the volume fraction
    has been properly accounted for.

  \item Using twice the intrinsic viscosity of a tank treading vesicle
    was incorrect.  We now compare the intrinsic viscosity of a single
    tank treading vesicle with the doublet.

  \item The total force $\xxi$ is the sum of a bending, tension, and
    adhesion force.  Therefore, the contribution to the intrinsic
    viscosity due to each of these forces can be individually computed.
\end{itemize}
